---
title: "R Notebook"
output: html_notebook
---

My personal project
Analysing personal budget

Questions:

Is it cheaper to live in Scotland than in England?
Have we saved money since the start of the pandemic?
What is the difference between living costs in England and in Scotland?

Reading libraries
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

Reading in data
```{r}
budget <- read_csv("Joint Budget as of 2022-02-22 0208 PM - Budget.csv")
register <- read_csv("Joint Budget as of 2022-02-22 0208 PM - Register.csv")
```

initial data exploration
```{r}
head(budget)
```

Notes for initial data cleaning:
- dates need to be cleaned
- column names need to be cleaned
- some variables likely need to be fixed (there seem to be quite a lot)
- not clear what the difference is between budget and register
- I'll initially focus on the "budget" dataset - the register seems confusing?

```{r}
budget_clean <- budget %>%
  clean_names() %>%
  separate(month, into = c("month", "year"), sep = " ", remove = TRUE) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(month = as.numeric(case_when(month == "Jan" ~ "01",
                           month == "Feb" ~ "02",
                           month == "Mar" ~ "03",
                           month == "Apr" ~ "04",
                           month == "May" ~ "05",
                           month == "Jun" ~ "06",
                           month == "Jul" ~ "07",
                           month == "Aug" ~ "08",
                           month == "Sep" ~ "09",
                           month == "Oct" ~ "10",
                           month == "Nov" ~ "11",
                           month == "Dec" ~ "12")))
  # mutate(date = as.numeric(date))
  
```

```{r}
  budget_clean %>%
  distinct(month)                        
```


```{r}
budget_clean %>%
  distinct(category_group)
```

```{r}
budget_clean %>%
  distinct(category_group_category)
```
So this is problematic - why are there two columns with the same function? From an overview glance, it seems like cat_grp_cat can be dropped, as it combines the other two.


```{r}
budget_clean <- budget_clean %>%
  select(-category_group_category)
```

Okay, so then looking at the other two...

```{r}
budget_clean %>%
  distinct(category_group)
```

```{r}
budget_clean %>%
  distinct(category) %>%
  arrange(category)
```
That's a LOT... 140. And I can definitely see doubles (some are mistakes). So I will clean this up.

SO - I will:
- rename the columns (main, and subcat)
- change the subcat variables

```{r}
# renaming the columns
budget_clean <- budget_clean %>%
  rename(main_category = "category_group",
         sub_category = "category")
```

```{r}
budget_clean %>%
  mutate(sub_category = case_when(sub_category == "⚠️ Emergency Buffer" ~ " ",
                                  sub_category == "Babysitting" == " " ~ " ",
                                  sub_category == "Back Extension" == " " ~ " ",
                                  sub_category == "Barry" == " " ~ " ",
                                  sub_category == "Barry from us" == " " ~ " ",
                                  sub_category == "Bathroom" == " " ~ " ",
                                  sub_category == "BIG Holiday 2026ish" == " " ~ " ",
                                  sub_category == "Bill and Rae" == " " ~ " ",
                                  sub_category == "Billy and Rae	" == " " ~ " ",
                                  sub_category == "Birthday Bike" == " " ~ " ",
                                  sub_category == "Books" == " " ~ " ",
                                  sub_category == "Car Breakdown (Annual)" == " " ~ " ",
                                  sub_category == "Car Finance" == " " ~ " ",
                                  sub_category == "Car Insurance (Annual)" == " " ~ " ",
                                  sub_category == "Car Insurance (Monthly)" == " " ~ " ",
                                  sub_category == "Car MOT (Annual)" == " " ~ " ",
                                  sub_category == "Car Seat" == " " ~ " ",
                                  sub_category == "Car Services" == " " ~ " ",
                                  sub_category == "Car Tax (Annual)" == " " ~ " ",
                                  sub_category == "Cards/Presents" == " " ~ " ",
                                  sub_category == "Caroline and Richard" == " " ~ " ",
                                  sub_category == "Caroline, Richard" == " " ~ " ",
                                  sub_category == "Charlie" == " " ~ " ",
                                  sub_category == "Chloe" == " " ~ " ",
                                  sub_category == "Chloe Clothes" == " " ~ " ",
                                  sub_category == "Chloe Essentials" == " " ~ " ",
                                  sub_category == "Chloe from us" == " " ~ " ",
                                  sub_category == "Chloe Toys/Books" == " " ~ " ",
                                  sub_category == "Chloe's Bedroom" == " " ~ " ",
                                  sub_category == "Chloe's Birthday" == " " ~ " ",
                                  sub_category == "Chloe's Future	" == " " ~ " ",
                                  sub_category == "Christmas 2021 Pot" == " " ~ " ",
                                  sub_category == "Christmas 2022 Pot" == " " ~ " ",
                                  sub_category == "Cinema" == " " ~ " ",
                                  sub_category == "Clothing" == " " ~ " ",
                                  sub_category == "Contact Lenses" == " " ~ " ",
                                  sub_category == "Council Tax" ~ " ",
                                  sub_category == "Crafts" ~ " ",
                                  sub_category == "DBS Renewal Fee" ~ " ",
                                  sub_category == "Dental" ~ " ",
                                  sub_category == "Disney+ (Annual)" ~ " ",
                                  sub_category == "Eating Out / Take Out" ~ " ",
                                  sub_category == "Ed and Ling" ~ " ",
                                  sub_category == "Ed, Ling" ~ " ",
                                  sub_category == "Energy" ~ " ",
                                  sub_category == "Entertainment" ~ " ",
                                  sub_category == "Ezra" ~ " ",
                                  sub_category == "Factor Fees (3-5 months)" ~ " ",
                                  sub_category == "Family Holiday 2022" ~ " ",
                                  sub_category == "Football" ~ " ",
                                  sub_category == "Front Door Loan Repayment" ~ " ",
                                  sub_category == "Gaming Upgrade" ~ " ",
                                  sub_category == "Garden" ~ " ",
                                  sub_category == "Gardening Disposal" ~ " ",
                                  sub_category == "Giving" ~ " ",
                                  sub_category == "Grandad from Chloe" ~ " ",
                                  sub_category == "Gray (NCT)" ~ " ",
                                  sub_category == "Gym" ~ " ",
                                  sub_category == "Halifax CC" ~ " ",
                                  sub_category == "HCPC Fee" ~ " ",
                                  sub_category == "Home Decoration" ~ " ",
                                  sub_category == "Home Insurance (Annual)" ~ " ",
                                  sub_category == "Home Security" ~ " ",
                                  sub_category == "House Pot" ~ " ",
                                  sub_category == "Internet and TV" ~ " ",
                                  sub_category == "Isla" ~ " ",
                                  sub_category == "Joe" ~ " ",
                                  sub_category == "Joe from Lou" ~ " ",
                                  sub_category == "Joe Parents	" ~ " ",
                                  sub_category == "Joe's Phone" ~ " ",
                                  sub_category == "Kirsten" ~ " ",
                                  sub_category == "Kirsten from us" ~ " ",
                                  sub_category == "Kirsten, Søren" ~ " ",
                                  sub_category == "Kitchen" ~ " ",
                                  sub_category == "Leaving Do" ~ " ",
                                  sub_category == "Lee and Ola" ~ " ",
                                  sub_category == "Life Insurance" ~ " ",
                                  sub_category == "Living Room" ~ " ",
                                  sub_category == "Loft" ~ " ",
                                  sub_category == "Lou" ~ " ",
                                  sub_category == "Lou from Joe" ~ " ",
                                  sub_category == "Lou Parents" ~ " ",
                                  sub_category == "Lou's Birthday Clothes" ~ " ",
                                  sub_category == "Lou's Birthday Fund 2020" ~ " ",
                                  sub_category == "Lou's Future" ~ " ",
                                  sub_category == "Lou's iPhone" ~ " ",
                                  sub_category == "Lou's Phone" ~ " ",
                                  sub_category == "Louisa, James, Ciaran" ~ " ",
                                  sub_category == "Lysander" ~ " ",
                                  sub_category == "Matt and Anita" ~ " ",
                                  sub_category == "Medical" ~ " ",
                                  sub_category == "Morfar from Chloe" ~ " ",
                                  sub_category == "Mormor from Chloe" ~ " ",
                                  sub_category == "Mortgage" ~ " ",
                                  sub_category == "Mortgage Life Cover" ~ " ",
                                  sub_category == "Moving Costs" ~ " ",
                                  sub_category == "Nana's Shoe Fund for Chloe" ~ " ",
                                  sub_category == "Nanna from Chloe" ~ " ",
                                  sub_category == "Nectar CC" ~ " ",
                                  sub_category == "Nete" ~ " ",
                                  sub_category == "New Car Deposit" ~ " ",
                                  sub_category == "New Glasses" ~ " ",
                                  sub_category == "New Laptop for Joe" ~ " ",
                                  sub_category == "New Laptop for Lou" ~ " ",
                                  sub_category == "New TV" ~ " ",
                                  sub_category == "Nursery/Childcare" ~ " ",
                                  sub_category == "Olee" ~ " ",
                                  sub_category == "Online Services" ~ " ",
                                  sub_category == "Online Storage" ~ " ",
                                  sub_category == "Other" ~ " ",
                                  sub_category == "Our Bedroom" ~ " ",
                                  sub_category == "Owed To Us" ~ " ",
                                  sub_category == "Pakkeleg" ~ " ",
                                  sub_category == "Parking Permit" ~ " ",
                                  sub_category == "Petrol" ~ " ",
                                  sub_category == "Rachel" ~ " ",
                                  sub_category == "Rachel from us" ~ " ",
                                  sub_category == "Rachel, Barry" ~ " ",
                                  sub_category == "Relocation Expenses" ~ " ",
                                  sub_category == "Rent" ~ " ",
                                  sub_category == "Services" ~ " ",
                                  sub_category == "Social Workers Union" ~ " ",
                                  sub_category == "Søren" ~ " ",
                                  sub_category == "Soren from us" ~ " ",
                                  sub_category == "Spending Money" ~ " ",
                                  sub_category == "Stocking fillers" ~ " ",
                                  sub_category == "Stocking Fillers" ~ " ",
                                  sub_category == "Study" ~ " ",
                                  sub_category == "Therapy" ~ " ",
                                  sub_category == "Transportation" ~ " ",
                                  sub_category == "Trips" ~ " ",
                                  sub_category == "TV License (Annual)" ~ " ",
                                  sub_category == "Upstairs Hall" ~ " ",
                                  sub_category == "Visitor Permits" ~ " ",
                                  sub_category == "Water" ~ " ",
                                  sub_category == "Weekly Food" ~ " ",
                                  sub_category == "Work Expenses" ~ " ",
                                  sub_category == "Workspace Upgrades" ~ " ",
                                  sub_category == "YNAB (Annual)" ~ " "
                                  ))
```

```{r}
budget_clean %>%
  distinct(sub_category) %>%
  arrange(sub_category)
```

```{r}
budget_clean %>%
  filter(sub_category == "Weekly Food" & year <= 2020 & month <= 9) %>%
  ggplot(aes(x = year, y = activity, fill = month)) +
  geom_col() +
  facet_wrap(~month)

budget_clean %>%
  filter(sub_category == "Weekly Food" & year >= 2020 & month >= 9) %>%
  ggplot(aes(x = year, y = activity, fill = month)) +
  geom_col() +
  facet_wrap(~month)

```


