---
title: "R Notebook"
output: html_notebook
---

My personal project
Analysing personal budget

Questions:

Is it cheaper to live in Scotland than in England?
Have we saved money since the start of the pandemic?
What is the difference between living costs in England and in Scotland?

Reading libraries
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

Reading in data
```{r}
budget <- read_csv("Joint Budget as of 2022-02-22 0208 PM - Budget.csv")
register <- read_csv("Joint Budget as of 2022-02-22 0208 PM - Register.csv")
```

initial data exploration
```{r}
head(budget)
```

Notes for initial data cleaning:
- dates need to be cleaned
- column names need to be cleaned
- some variables likely need to be fixed (there seem to be quite a lot)
- not clear what the difference is between budget and register
- I'll initially focus on the "budget" dataset - the register seems confusing?

```{r}
budget %>%
  clean_names() %>%
  separate(month, into = c("month", "year"), sep = " ", remove = TRUE) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(month = as.numeric(case_when("Apr" == "Apr" ~ "04",
                            "May" == "May" ~ "05",
                            "Jun" == "Jun"  ~ "06",
                            "Jul" == "Jul"  ~ "07",
                            "Aug" == "Aug"  ~ "08",
                            "Sep" == "Sep"  ~ "09",
                            "Oct" == "Oct"  ~ "10",
                            "Nov" == "Nov"  ~ "11",
                            "Dec" == "Dec"  ~ "12",
                            "Jan" == "Jan"  ~ "01",
                            "Feb" == "Feb"  ~ "02",
                            "Mar" == "Mar"  ~ "03"))) %>%
  
  
```


```{r}
budget %>%
  ggplot(aes(x = Month, y = Activity)) +
  geom_col() 
```

